---
title: "Home"
output:  
    html_document:  
        toc: true  
        toc_float: true 
        collapsed: false
        smooth_scroll: false
        number_sections: true
        df_print: paged
        code_folding: hide
---

# Rmarkdown website for easy introduction to R

### Use the Navigation bar to see different markdown with examples

## Some interesting links:

- [Data carpentry](https://datacarpentry.org/)
- [R for Data Science](https://r4ds.had.co.nz/)
- [Tidyverse cookbook](https://rstudio-education.github.io/tidyverse-cookbook/)

# Basics of R

You can do simple operations and assing that to variables
```{r, echo=TRUE}
variable_number <- 2+2

variable_number
```
You could also have vectors which represent lists of numbers or characters.
You can create a vector by concatenating values c()
```{r, echo=TRUE}
vector_numbers <- c(variable_number, 2, 4, 5)

vector_characters <- c("Reactor", "pH", "TOC", "TN")
```

When you join them you can generate dataframes, which combine characters, factors and numbers
```{r, echo=TRUE}
first_column <- rnorm(30, 3, 0.2)
second_column <- rnorm(30, 8, 1.2)
third_column <- runif(30, 1, 8)
dataframe <- data.frame(first_column, second_column, third_column)
head(dataframe, 5)
```

We can change the name to the dataframe columns
```{r, echo=TRUE}
names(dataframe) <- c("one", "two", "three")
head(dataframe, 5)
```


Now we could select specific rows or columns by the bracket operator 

dataframe[rows, columns]
```{r, echo=TRUE}
dataframe[1:4,1:2]
```

we could even plot one against each other

```{r, echo=TRUE}
plot(dataframe$one, dataframe$two)
```

Plots have many possible arguments, for example:

```{r, echo=TRUE}
plot(dataframe$one, dataframe$two, 
     xlab = "x axis", ylab="y axis", main="A simple plot", 
     las =1,
     pch =19, col = "blue")

```

